Discrete Quantum Theories and Computing

### 2019-02-20 dissertation.tex
dissertation.lyx "6.1 Gleason's Theorem for General QIVPMs" Remove the last TODO... CV_thesis.pdf Remove "(expect)"...

\section{Discrete Quantum Theory (I): Irreducible Discrete $D$-dimensional
States\label{DQCnqubitBloch.sec}}

@@

\section{Discrete Quantum Theory (I): Geometry of Entangled States\label{sec:Geometry-of-Entangled-States}}

\begin{figure}[!b]
\begin{centering}
\includegraphics[width=1\columnwidth]{statePlot3}
\par\end{centering}
\caption{\label{statePlot3Unit.fig}Logarithmic plot of the number of discrete
unnormalized states (top, in red), vs the number of normalized discrete
states (middle, in blue), vs the irreducible states (bottom, in green)
for the first 6 $\Fpp$-compatible primes, ($3$, $7$, $11$, $19$,
$23$, $31$), for the number of qubits $1$, $2$, $3$, and $4$.}
\end{figure}
To discuss entanglement, we consider a $D$-dimensional quantum system
composed of $n$-qubit subsystems, i.e., $D=2^{n}$ as usual. Without
regard to uniqueness, an $n$-qubit state with discrete complex coefficients
in $\Fpp$ will have the total possible space of coefficients with
dimension $p^{2\times2^{n}}$ (including the null state). Imposing
the condition of a length-one norm in $\Fp$, this number is reduced
to $p^{2^{n}-1}\left(p^{2^{n}}-1\right)$. The ratio of all the states
to the unit-norm states is asymptotically $p$: 
\begin{equation}
\frac{p^{2^{n}+1}}{p^{2^{n}}-1}\rightarrow p\,,
\end{equation}
so there are roughly $p$ sets of coefficients, for any number of
qubits $n$, that are discarded for each retained unit-length state
vector. A factor of $p+1$ more states is discarded in forming the
discrete Bloch sphere of irreducible states. Selected plots of the
full space compared to both the unit-norm space and the irreducible
space for a selection of complexified finite fields are shown in Figure~\ref{statePlot3Unit.fig}
for 1, 2, 3, and 4 qubits.

@@

With different settings, whether there is an approximate version of
Gleason's theorem was asked by Sam Sanders in constructivenews on
2013~\cite{Sanders2013}, and there is no clear answer for his question
as far as I understand. To have an idea of how hard this question
could be, recall in Sec.~\ref{sec:fuzzy} we state that a quantum
probability space is glued by a family of classical probability spaces.
\todo{Check the differential geometry analog with Andy, since I don't
really understand the differential geometry\ldots{} }This is like
the situation that a manifold is glued by many local coordinates.
When each small piece has \emph{exactly} the same and positive curvature,
the Killing-Hopf theorem asserts this manifold is a sphere, but little
can we say even if the curvature has a small deviation from constant.
A similar situation might happen when approximating Gleason's theorem,
but this time the whole space is glued by ``local'' classical probability
space defined by each orthonormal basis $\left\{ \ket{\psi_{i}}\right\} _{i=0}^{2}$.
When the sum of $\frameL$, $\sum_{i=0}^{2}\frameL\left(\ket{\psi_{i}}\right)$,
is \emph{exactly} the same and equal to $1$, Gleason's theorem asserts
that $\frameL$ can be expressed as Eq.~(\ref{eq:Gleason-for-CP2}).
However, when each local classical probability space becomes imprecise,
a general $\frameL$ might be as wild as we can imagine, and we might
need to know a bit more, like its QIVPM is a composite function, to
deduce its global property.

>>>>

\section{Irreducible Discrete $D$-dimensional States\label{DQCnqubitBloch.sec}}

@@

\section{Geometry of Entangled States\label{sec:Geometry-of-Entangled-States}}

To discuss entanglement, we consider a $D$-dimensional quantum system
composed of $n$-qubit subsystems, i.e., $D=2^{n}$ as usual. Without
regard to uniqueness, an $n$-qubit state with discrete complex coefficients
in $\Fpp$ will have the total possible space of coefficients with
dimension $p^{2\times2^{n}}$ (including the null state). Imposing
the condition of a length-one norm in $\Fp$, this number is reduced
to $p^{2^{n}-1}\left(p^{2^{n}}-1\right)$. The ratio of all the states
to the unit-norm states is asymptotically $p$: 
\begin{equation}
\frac{p^{2^{n}+1}}{p^{2^{n}}-1}\rightarrow p\,,
\end{equation}
so there are roughly $p$ sets of coefficients, for any number of
qubits $n$, that are discarded for each retained unit-length state
vector. A factor of $p+1$ more states is discarded in forming the
discrete Bloch sphere of irreducible states. Selected plots of the
full space compared to both the unit-norm space and the irreducible
space for a selection of complexified finite fields are shown in Figure~\ref{statePlot3Unit.fig}
for 1, 2, 3, and 4 qubits.
\begin{figure}[!b]
\begin{centering}
\includegraphics[width=1\columnwidth]{statePlot3}
\par\end{centering}
\caption{\label{statePlot3Unit.fig}Logarithmic plot of the number of discrete
unnormalized states (top, in red), vs the number of normalized discrete
states (middle, in blue), vs the irreducible states (bottom, in green)
for the first 6 $\Fpp$-compatible primes, ($3$, $7$, $11$, $19$,
$23$, $31$), for the number of qubits $1$, $2$, $3$, and $4$.}
\end{figure}

@@

With different settings, whether there is an approximate version of
Gleason's theorem was asked by Sam Sanders in constructivenews on
2013~\cite{Sanders2013}, and there is no clear answer for his question
as far as I understand. To have an idea of how hard this question
could be, recall in Sec.~\ref{sec:fuzzy} we state that a quantum
probability space is glued by a family of classical probability spaces.
This is like the situation that a manifold is glued by many local
coordinates. When each small piece has \emph{exactly} the same and
positive curvature, the Killing-Hopf theorem asserts this manifold
is a sphere, but little can we say even if the curvature has a small
deviation from constant. A similar situation might happen when approximating
Gleason's theorem, but this time the whole space is glued by ``local''
classical probability space defined by each orthonormal basis $\left\{ \ket{\psi_{i}}\right\} _{i=0}^{2}$.
When the sum of $\frameL$, $\sum_{i=0}^{2}\frameL\left(\ket{\psi_{i}}\right)$,
is \emph{exactly} the same and equal to $1$, Gleason's theorem asserts
that $\frameL$ can be expressed as Eq.~(\ref{eq:Gleason-for-CP2}).
However, when each local classical probability space becomes imprecise,
a general $\frameL$ might be as wild as we can imagine, and we might
need to know a bit more, like its QIVPM is a composite function, to
deduce its global property.
